---
title: "Forest clearance data (prior to 2014)"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

There are three types of clearance data on the PARIVESH portal, hosted by the MOEF&CC. These include: Environmental, Forests and Wildlife. The data for Forests and Wildlife are quantifiable, although the data for Wildlife was not available as an excel file and this information need to be obtained from the website. 

All analysis being performed is through the R programming environment. Please feel free to edit and use this code as you please. The data used for this purpose is available through the 'Data' folder. 

```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(stringr)
library(sf)
library(lubridate)
library(scico)
library(ggrepel)
```


First, we will carry out some exploratory data analysis using the Forest clearance data. The data available has been stored in 5 separate excel files:

1. All data prior to 2014
2. Data after 2014 is subdivided into:
    a) Allocation of fresh forest land (Form-A)	
    b) Application Under Section 2(iii)	
    c) Renewal of lease (Form-B)	
    d) Prospecting of Minerals (Form-C)


Let's first analyze the data prior to 2014

```{r}

## 1. Let's read in the data prior to 2014
pre_2014 <- read.csv("Data/forest_clearance/for_clear_before_2014.csv",
                     stringsAsFactors = F)
head(pre_2014) # Take a look at the first 6 records
names(pre_2014)

# Shall subset certain categories
pre_2014 <- pre_2014[,c(4,5,7,8,9,10)]

# We have 6 categories: Proposal Name, Category the proposal falls under, Area applied for, Date (around which it was approved?), State Name and Proposal Status

# Convert the date category from a factor to date
names(pre_2014)[4] <- "DATE"
pre_2014$DATE <- year(dmy(pre_2014$DATE)) # Keeping only years for now

```


A couple of analyses and plots for the forest clearance data:  

The total number of forest clearances prior to 2014 that were approved and/or in other categories including pending etc was 31,883  

The total number of forest clearances prior to 2014 that were rejected / withdrawn / closed was 5,768  

Let's summarize the area of forests cleared vs. those not cleared (Rejected proposals)  

21,63,215 hectares of forests were approved to be cleared prior to 2014/pending   
8,48,475 hectares of forests were not cleared or proposals were rejected/withdrawn  

In other words, over 84.68% of proposals submitted were approved or pending to be approved! 

```{r}

# Area saved / rejected proposals
n_reject <- pre_2014 %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>%
  filter(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN") %>% summarise() %>% tally(AREA_APPLIED)%>%
  tally(n)

n_reject # 8,48,475 hectares of forests were not cleared prior to 2014

# Area cleared / approved proposals and/or pending
n_appro <- pre_2014 %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>% summarise() %>%   tally(AREA_APPLIED) %>% tally(n) - n_reject

n_appro # 21,63,215 hectares of forests were cleared or approved to be cleared / pending

```

Let's calculate the area of forests approved or pending to be approved as function of the category they were binned under:

Encroachments and mining contributed most to the loss of forest area prior to 2014.

```{r}

# Here onwards we will use only the proposals that were approved or pending to be approved
appro_pre_2014 <- pre_2014 %>%
  filter(!(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN")) %>%
  filter(!DATE=="1882")

# Data by category and area
cat_area <- appro_pre_2014 %>% 
  group_by(CATEGORY) %>%
  summarise(Total_area = sum(AREA_APPLIED)) %>% 
  ungroup() %>% 
  mutate(CATEGORY = stringr::str_to_title(CATEGORY))

# Create a plot
cat_plot <-
ggplot(cat_area %>% 
         filter(Total_area > 10), 
       aes(x = CATEGORY , y = Total_area, 
           fill= CATEGORY)) +   
  geom_bar(stat = "identity",
           color = "black", 
           size = 0.08,
           position = position_dodge()) +
  geom_text(aes(label = scales::comma(round(Total_area))), 
            position = position_dodge(width=1),
            size = 3, vjust = -0.5)+ 
  theme_minimal(base_family = "TT Arial") +
  scale_y_continuous(trans = "sqrt",
                     labels = scales::comma,
                     breaks = c(0.25, 0.5, 1:6) * 1e5) +
  scale_x_discrete(guide = guide_axis(n.dodge = 4)) +
  labs(x = "Category", 
       y = "Total Area (in hectares)") +
  theme(#axis.text.x = element_text(angle=, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none",
        panel.grid.minor =  element_blank()) +
  scico::scale_fill_scico_d(palette = "roma")

ggsave(cat_plot, filename = "Figures/cat_area_pre2014.png",
       width= 12, height= 7, dpi = 300)

```


Calculating area of forests approved to be cleared or pending to be cleared per year

See years 1984, 1989, 1995 and 2006

```{r}
# Data by year and area
appro_pre_2014$DATE <- as.character(appro_pre_2014$DATE)

area_year <- appro_pre_2014 %>% group_by(DATE) %>%
  summarise(Total_area = sum(AREA_APPLIED))

# Create a plot
area_year_plot <-
ggplot(area_year, 
       aes(x = as.numeric(DATE) , y = Total_area)) +   
  geom_line(stat = "identity", #color = "black", 
           size = 0.3, #position = position_dodge(), 
           col = "blue", alpha = 1) +
  geom_point(stat = "identity", #color = "black", 
           size = 2, #position = position_dodge(), 
           col = "blue", alpha = 1) +
  
  geom_text(aes(label = scales::comma(round(Total_area))), 
            position = position_dodge(width=1),
            size = 3, vjust = -0.5) + 
  
  scale_x_continuous(breaks = c(1975:2015),
                     guide = guide_axis(n.dodge = 2))+
  scale_y_continuous(trans = "sqrt",
                     labels = scales::comma,
                     breaks = c(0.05, 0.5, 1:3) * 1e5) +
  
  theme_minimal(base_family = "TT Arial") +
  labs(x="Year", y="Total Area (in hectares)") +
  theme(# axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none")

area_year_plot

ggsave(area_year_plot, filename = "Figures//year_area_pre2014.png",
       width=12, height=7,dpi = 300)

```

Lets calculate the area of forests approved or pending to be approved as function of the state.  

```{r}
library(sf)
# read in states
states <- st_read("Data/spatial/indian_states.gpkg") %>% 
  mutate(state_name = as.character(state_name),
         state_name = case_when(
    state_name == "Jammu & Kashmir" ~ "Jammu And Kashmir",
    state_name == "Arunanchal Pradesh" ~ "Arunachal Pradesh",
    state_name == "Andaman & Nicobar Island" ~ "Andaman And Nicobar",
    T ~ state_name
  ))

state_area <- appro_pre_2014 %>% 
  group_by(STATE_NAME) %>%
  summarise(Total_area = sum(AREA_APPLIED)) %>% 
  ungroup() %>% 
  mutate(STATE_NAME = stringr::str_to_title(STATE_NAME)) %>% 
  mutate(STATE_NAME = case_when(
    STATE_NAME == "Dadar & Nagar Haveli" ~ "Dadara & Nagar Haveli",
    STATE_NAME == "Pondicherry" ~ "Puducherry",
    STATE_NAME == "Orissa" ~ "Odisha",
    STATE_NAME == "Delhi" ~ "NCT of Delhi",
    T ~ STATE_NAME))

# join spatial with area
states <- states %>% 
  left_join(state_area,
            by = c("state_name" = "STATE_NAME")) %>% 
  bind_cols(as_tibble(st_coordinates(st_centroid(states))))

# Create a plot
state_area_plot <-
ggplot(states) +
  geom_sf(aes(fill = Total_area),
          col = "black",
          size = 0.1)+
  geom_label_repel(aes(X, Y,
                 label = scales::comma(as.integer(Total_area),
                                       accuracy = 1)),
             size = 2,
             col = "black",
             alpha = 0.8)+
  scale_fill_scico(palette = "hawaii",
                       direction = -1,
                   na.value = "black",
                       trans = "sqrt",
                   name = "Total area",
                   breaks = c(0.1, 0.5, 1, 2, 4, 6) * 1e5,
                   labels = scales::comma)+
  
  ggthemes::theme_map(base_family = "TT Arial") +
  labs(title = "Total area cleared pre-2014")

ggsave(state_area_plot, 
       filename = "Figures/state_area_pre2014.png",
       width=12, height=7,dpi = 300)

```


Lets calculate the area of forests approved or pending to be approved as function of the state and the category. 

Madhya Pradesh lost a vast chunk of it's forests prior to 2014, followed by Arunachal Pradesh - two states that possess an astounding level of biodiversity

```{r}
# group by category and state
state_area <- appro_pre_2014 %>% group_by(CATEGORY, STATE_NAME) %>%
  summarise(Total_area=sum(AREA_APPLIED))

nb.cols <- 23
mycolors <- colorRampPalette(brewer.pal(11, "Paired"))(nb.cols)

# state_plot <- 
ggplot(state_area, aes(x = CATEGORY , y = Total_area, fill= CATEGORY)) +   
  geom_bar(stat="identity",color="black",size=0.08, position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=1.8, vjust=-0.5) + theme_minimal() +
  facet_wrap(~STATE_NAME) +
  labs(x="Category", y="Total Area (in hectares)") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(size=8), 
        legend.text=element_text(size=8)) +
  scale_fill_manual(values = mycolors)
 
state_plot

ggsave(state_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\state_cat_area_pre2014.png",
       width=20, height=10,dpi = 300)
```


Lets calculate the area of forests approved or pending to be approved as function of the category and year

Encroachments and mining - 1989, 2005 and 1995

```{r}

cat_year <- appro_pre_2014 %>% group_by(CATEGORY, DATE) %>% 
  summarise(Total_area = sum(AREA_APPLIED))

cat_year_plot <- ggplot(cat_year, aes(x = CATEGORY , y = Total_area, fill= CATEGORY)) + 
  geom_bar(stat="identity",color="black",size=0.08, position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=1.8, vjust=-0.5) + theme_minimal() +
  facet_wrap(~DATE) +
  labs(x="Year", y="Total Area (in hectares)") + scale_fill_manual(values = mycolors) +
   theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(size=5), 
        legend.text=element_text(size=5))

cat_year_plot

ggsave(cat_year_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\cat_year_pre2014.png",
       width=20, height=10,dpi = 300)
  

```
