---
title: "Analysis of 3 time periods: 2000-2006; 2006-2014 and 2014 to Present"
output: html_notebook
---

There are three types of clearance data on the PARIVESH portal, hosted by the MOEF&CC. These include: Environmental, Forests and Wildlife. The data for Forests and Wildlife are quantifiable, although the data for Wildlife was not available as an excel file and this information had to be scraped from their website. 

All analysis being performed is through the R programming environment. Please feel free to edit and use this code as you please. The data used for this purpose is available through the 'Data' folder. 

```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(lubridate)

```


First, we will carry out some exploratory data analysis using the Forest clearance data. The data available has been stored in 5 separate excel files:

1. All data prior to 2014
2. Data after 2014 is subdivided into:
    a) Allocation of fresh forest land (Form-A)	
    b) Application Under Section 2(iii)	
    c) Renewal of lease (Form-B)	
    d) Prospecting of Minerals (Form-C)


Let's first analyze data as binned under 3 time periods:

1. 2000 to 2006: Prior to the 2006 amendment of the EIA
2. 2006 to 2014: After the 2006 amendment of the EIA
3. 2014 to Present: An artifical divide as mentioned by the MOEF&CC

```{r}
## 1. Let's read in the data prior to 2014
pre_2014 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_before_2014.csv",stringsAsFactors = F)
head(pre_2014) # Take a look at the first 6 records
names(pre_2014)

# Shall subset certain categories
pre_2014 <- pre_2014[,c(4,5,7,8,9,10)]

# We have 6 categories: Proposal Name, Category the proposal falls under, Area applied for, Date (around which it was approved?), State Name and Proposal Status

# Convert the date category from a factor to date
names(pre_2014)[4] <- "DATE"
pre_2014$DATE <- year(dmy(pre_2014$DATE)) # Keeping only years for now

## 1. Let's read in the .csv's for data post 2014
dat1 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clearance_forestLand_after_2014.csv",stringsAsFactors = F)
dat2 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_form2_after_2014.csv",stringsAsFactors = F)
dat3 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_leaseRenew_after_2014.csv",stringsAsFactors = F)
dat4 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_miningProspect_after_2014.csv",stringsAsFactors = F)

# Subset necessary columns
dat1 <- dat1[,c(2,6,7,9,10,11)]
dat2 <- dat2[,c(2,6,7,9,10,11)]
dat3 <- dat3[,c(10,5,6,8,9,11)]
dat4 <- dat4[,c(10,5,6,8,9,11)]

names(dat1); names(dat2); names(dat3); names(dat4)

names(dat3)[1] <- "state_name"
names(dat4)[1] <- "state_name"

post_2014 <- bind_rows(dat1,dat2,dat3,dat4)

# We have 6 categories: Proposal Name, category the proposal falls under, Area applied for, date (around which it was approved?), State Name and Proposal Status

# Convert the date category from a factor to date
names(post_2014)[5] <- "date"
post_2014$date <- year(dmy(post_2014$date)) # Keeping only years for now

```

Subsetting the data
```{r}
dat_2000_2006 <- pre_2014 %>% filter(DATE>="2000") %>%
  filter(DATE<="2006")

dat_2006_2014 <- pre_2014 %>% filter(DATE>="2006") %>%
  filter(DATE<="2014")

```

Summary of number of grants approved and rejected across these time periods

Between 2000 and 2006, 5076 proposals were categorized as approved / under one of the pending categories. During the same period, 1213 proposals were rejected. 
Rejection Rate: 19.3%

Between 2006 and 2014, 10815 proposals were categorized as approved / under one of the pending categories. During the same period, 1381 proposals were rejected. 
Rejection rate: 11.3%

Between 2014 and 2020, 17037 proposals were categorized as approved / under one of the pending categories. During the same period, 120 proposals were rejected. 
Rejection rate: 0.7%!


```{r}

# Replace dat_2006_2014 with dat_2000_2006 for the other time period.
n_reject <- dat_2006_2014 %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>%
  filter(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN") %>% summarise() 

n_appro <- dat_2006_2014 %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>% summarise()

# For post 2014:
n_reject_2014 <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>%
  filter(proposal_status=="Returned" | proposal_status=="Rejected" |
         proposal_status=="Closed" | proposal_status=="Rejected By REC" |
           proposal_status=="Withdrawn" | proposal_status=="CLOSED" |proposal_status=="REVOKED") %>% summarise()

n_reject_2014 

# Area cleared / approved proposals and/or pending
n_appro_2014 <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>% summarise() 

n_appro_2014

```

Forest area in hectares stated to be cleared / categorized as pending approval

2000 to 2006: Area stated to be cleared / potentially cleared - 6,29,638 hectares
2000 to 2006: Area saved - 254555.6 hectares

2006 to 2014: Area stated to be cleared / potentially cleared - 7,51,354.8 hectares
2006 to 2014: Area saved - 45302.78 hectares

2014 to Present: Area stated to be cleared / potentially cleared - 14,82,247 hectares
2014 to Present: Area saved - 13077.03 hectares


```{r}

# Pre 2014 data

# Area saved / rejected proposals
n_reject <- dat_2006_2014 %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>%
  filter(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN") %>% summarise() %>% tally(AREA_APPLIED)%>%
  tally(n)

n_reject # 8,48,475 hectares of forests were not cleared prior to 2014

# Area cleared / approved proposals and/or pending
n_appro <- dat_2006_2014  %>% select(PROPOSAL_STATUS,AREA_APPLIED) %>%
  group_by(PROPOSAL_STATUS, AREA_APPLIED) %>% summarise() %>%   tally(AREA_APPLIED) %>% tally(n) - n_reject

n_appro # 21,63,215 hectares of forests were cleared or approved to be cleared / pending

# Post 2014 data

# Area saved / rejected proposals
n_reject <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>%
  filter(proposal_status=="Returned" | proposal_status=="Rejected" |
         proposal_status=="Closed" | proposal_status=="Rejected By REC" |
           proposal_status=="Withdrawn" | proposal_status=="CLOSED" |proposal_status=="REVOKED") %>% summarise() %>% tally(area_applied)%>%
  tally(n)

n_reject # 13,077.03 hectares of forests were not cleared post 2014

# Area cleared / approved proposals and/or pending
n_appro <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>% summarise() %>%   tally(area_applied) %>% tally(n) - n_reject

n_appro # 14,82,247 hectares of forests were cleared or approved to be cleared / pending



```

Let's subset only those proposals that were approved. 

```{r}
# Here onwards we will use only the proposals that were approved or pending to be approved
appro_2000_2006 <- dat_2000_2006 %>%
  filter(!(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN"))

appro_2006_2014<- dat_2006_2014 %>%
  filter(!(PROPOSAL_STATUS=="REJECTED" | PROPOSAL_STATUS=="CLOSED" |
         PROPOSAL_STATUS=="RETURNED" | PROPOSAL_STATUS=="REVOKED" |
           PROPOSAL_STATUS=="WITHDRAWN"))

```


```{r}
# Data by category and area
cat_area_2006_2014 <- appro_2006_2014 %>% group_by(CATEGORY) %>%
  summarise(Total_area = sum(AREA_APPLIED))

# Create a plot
cat_plot <- ggplot(cat_area_2006_2014, aes(x = CATEGORY , y = Total_area, fill= CATEGORY)) +   geom_bar(stat="identity",color="black", size=0.08,position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+ theme_minimal() +
  labs(x="Category", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none") +
  scico::scale_fill_scico_d(palette = "roma")

cat_plot

ggsave(cat_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\cat_area_2006-2014.png",
       width=12, height=7,dpi = 300)

```

Year and area

```{r}
# Data by year and area
appro_2000_2006$DATE <- as.character(appro_2000_2006$DATE)
appro_2006_2014$DATE <- as.character(appro_2006_2014$DATE)

area_year <- appro_2006_2014 %>% group_by(DATE) %>%
  summarise(Total_area = sum(AREA_APPLIED))

# Create a plot
area_year_plot <- ggplot(area_year, aes(x = DATE , y = Total_area)) +   
  geom_bar(stat="identity",color="black", size=0.08, position=position_dodge(), fill="blue",alpha=0.4) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+ theme_minimal() +
  labs(x="Year", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none")

area_year_plot

ggsave(area_year_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\year_area_2006-2014.png",
       width=12, height=7,dpi = 300)

```

Area and state
```{r}
state_area <- appro_2006_2014 %>% group_by(STATE_NAME) %>%
  summarise(Total_area = sum(AREA_APPLIED))

# Create a plot
state_area_plot <- ggplot(state_area, aes(x = STATE_NAME , y = Total_area, fill=STATE_NAME)) +geom_bar(stat="identity",color="black", size=0.08,position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+
  theme_minimal() +
  labs(x="State", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none") +
  scico::scale_fill_scico_d(palette = "roma")

state_area_plot

ggsave(state_area_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\state_area_2006-2014.png",
       width=12, height=7,dpi = 300)
```

