---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Composite analysis of forest clearances

## Load data

```{r}
# Load required libraries
library(tidyverse)
library(readxl)
library(sf)
library(rlang)
library(lubridate)
library(scico)
library(ggrepel)
```

### Pre-2014 data 

```{r}
## 1. Let's read in the data prior to 2014
pre_2014 <- read_csv("Data/forest_clearance/for_clear_before_2014.csv")

# rename columns
pre_2014 <- pre_2014 %>% 
  `colnames<-`(str_replace_all(str_to_lower(colnames(.)),
                               " ", "_"))

# get year of applications
pre_2014 <- mutate(pre_2014,
                   year = year(dmy(date_from_ua_to_nodal))) %>% 
  filter(year >= 1994)
```

### Post-2014 data

```{r}
# get the same data for post 2014 period
post_2014 <- read_csv("Data/forest_clearance/for_clearance_forestLand_after_2014.csv")
post_2014 <- post_2014 %>% 
  `colnames<-`(str_replace_all(str_to_lower(colnames(.)),
                               " ", "_"))

# get year
post_2014 <- mutate(post_2014,
                    year = year(dmy(date_from_ua_to_nodal)))
```

### Bind data

```{r}
# bind
forest_data <- bind_rows(pre_2014, post_2014)

# correct case
forest_data$category <- str_to_lower(forest_data$category)
```

## Basic analyses

### What is the status of applications?

```{r}
# what is the most common status
count(forest_data, proposal_status) %>% arrange(-n)

# what are the most common categories weighted by area
count(forest_data, category, wt = area_applied) %>% 
  arrange(-n)

# what are the top categories
top_categories <- count(forest_data, category, wt = area_applied) %>%
  arrange(-n) %>% 
  .$category %>% head(5)


# get broader categories
forest_data <- forest_data %>% 
  mutate(status = case_when(
    str_detect(proposal_status, 
               regex("(PENDING|RETURNED|SECRETARY|COMPLIANCE|Under examination|Draft)",
                     ignore_case = TRUE)) ~
      "pending",
    str_detect(proposal_status, regex("APPROVED|IN-PRINCIPLE",
                                      ignore_case = TRUE)) ~
      "successful",
    str_detect(proposal_status, regex("REJECTED|REVOKED|WITHDRAWN|CLOSED",
                                      ignore_case = TRUE)) ~
      "unsuccessful",
    T ~ "other"
  ))

# get period
forest_data <- forest_data %>% 
  mutate(period = case_when(
    year < 2006 ~ "2000 - 2005",
    year >= 2006 & year < 2014 ~ "2006 - 2013",
    year >= 2014 ~ "2014 - 2020",
    T ~ NA_character_
  )) %>% 
  drop_na(period)

# relevel factors
forest_data <- forest_data %>% 
  mutate(status = forcats::fct_relevel(status,
                                       "successful",
                                       "pending",
                                       "unsuccessful",
                                       "other"))

# what is the area applied for in general
forest_data %>% 
  # filter(status != "unsuccessful") %>% 
count(category, status,
      period,
      wt = area_applied) %>% 
  filter(category %in% top_categories) %>% 
  
  group_by(category) %>% 
  mutate(total_area = sum(n)) %>% 
  
  arrange(category, desc(status)) %>% 
  
  ggplot() +
  geom_col(aes(x = reorder(category, total_area), 
               y = n,
               fill = status,
               group = category),
           position = "stack")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2),
                   label = str_to_title)+
  scale_y_continuous(labels = scales::comma) +
  scale_fill_scico_d(palette = "batlow") +
  facet_grid(~period) +
  labs(x = "Project Category",
       y = "Forest area cleared (hectares)")

ggsave("Figures/fig_area_by_category.png")


forest_data %>% 
  ggplot() +
  stat_count(aes(x = year, fill = status,
               weight = area_applied),
           size = 5,
           geom = "bar", 
           position = "stack") +
  scale_x_continuous(guide = guide_axis(n.dodge = 2),
                   label = str_to_title,
                   breaks = 1994:2020)+
  scale_y_continuous(labels = scales::comma) +
  scale_fill_scico_d(palette = "batlow") +
  facet_grid(~period, scales = "free_x") +
  coord_cartesian(expand = F) +
  labs(x = "Year",
       y = "Forest area cleared (hectares)")

ggsave("Figures/fig_area_by_year.png")
```

## Rule change years and other years

### Prepare data for models

```{r}
data_rule_change <- forest_data %>%
  group_by(year) %>% 
  summarise(total_area = sum(area_applied),
            proportion_approved = length(status[status == "successful"]) 
            / length(status[status %in% c("successful", "unsuccessful")])) %>% 
  mutate(event = if_else(year %in% c(1995, 2006),
                         "rule change", "other"))
```

### Area in relation to rule change

```{r}
# fit a simple linear model for area
# quasipoisson for overdispersion
model_area <- glm(total_area ~ event, 
   data = data_rule_change,
   family = "quasipoisson")

# print model summary
summary(model_area)

# print means and sd
data_area_summary <- group_by(data_rule_change, event) %>% 
  mutate(total_area_km = total_area * 0.01) %>% 
  summarise_at(vars("total_area_km"), .funs = list(mean_area = mean, 
                                                sd_area = sd))

data_area_summary %>% 
  knitr::kable()

# print % difference relative to normal years
abs(diff(data_area_summary$mean_area)) / min(data_area_summary$mean_area)
```

In years in which EIA rules were changed (1995 and 2006), applications are made to clear `r scales::percent(abs(diff(data_area_summary$mean_area)) / min(data_area_summary$mean_area))` more forest land than in other years.

### Proportion of success in relation to rule change

```{r}
# add time to present as a linear predictor
data_rule_change <- data_rule_change %>% 
  mutate(time_to_present = 2020 - year)

# fit a simple glm with binomial family
model_approval <- glm(proportion_approved ~ event,
                      data = data_rule_change,
                      family = "binomial")

# print summary
summary(model_approval)
```
