---
title: "Forest clearance data post 2014"
output:
  html_document:
    df_print: paged
---

There are three types of clearance data on the PARIVESH portal, hosted by the MOEF&CC. These include: Environmental, Forests and Wildlife. The data for Forests and Wildlife are quantifiable, although the data for Wildlife was not available as an excel file and this information needs to be obtained from their website. 

All analysis being performed is through the R programming environment. Please feel free to edit and use this code as you please. The data used for this purpose is available through the 'Data' folder. 

```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(lubridate)

```


First, we will carry out some exploratory data analysis using the Forest clearance data. The data available has been stored in 5 separate excel files:

1. All data prior to 2014
2. Data after 2014 is subdivided into:
    a) Allocation of fresh forest land (Form-A)	
    b) Application Under Section 2(iii)	
    c) Renewal of lease (Form-B)	
    d) Prospecting of Minerals (Form-C)


In this script, we will analyze all data post 2014

```{r}

## 1. Let's read in the .csv's for data post 2014
dat1 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clearance_forestLand_after_2014.csv",stringsAsFactors = F)
dat2 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_form2_after_2014.csv",stringsAsFactors = F)
dat3 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_leaseRenew_after_2014.csv",stringsAsFactors = F)
dat4 <- read.csv("C:\\Users\\Vijay\\Desktop\\envt_clearance\\forest_clearance\\for_clear_miningProspect_after_2014.csv",stringsAsFactors = F)

# Subset necessary columns
dat1 <- dat1[,c(2,6,7,9,10,11)]
dat2 <- dat2[,c(2,6,7,9,10,11)]
dat3 <- dat3[,c(10,5,6,8,9,11)]
dat4 <- dat4[,c(10,5,6,8,9,11)]

names(dat1); names(dat2); names(dat3); names(dat4)

names(dat3)[1] <- "state_name"
names(dat4)[1] <- "state_name"

post_2014 <- bind_rows(dat1,dat2,dat3,dat4)

# We have 6 categories: Proposal Name, category the proposal falls under, Area applied for, date (around which it was approved?), State Name and Proposal Status

# Convert the date category from a factor to date
names(post_2014)[5] <- "date"
post_2014$date <- year(dmy(post_2014$date)) # Keeping only years for now

```


A couple of analyses and plots for the forest clearance data:  

The total number of forest clearances post 2014 that were approved and/or in other categories including pending etc was 17,037 

The total number of forest clearances post 2014 that were rejected / withdrawn / closed was 120  

Let's summarize the area of forests cleared vs. those not cleared (Rejected proposals)  

14,82,247 hectares of forests were approved to be cleared/pending   
13,077.03 hectares of forests were not cleared or proposals were rejected/withdrawn  

In other words, 99.3% of proposals submitted were approved or pending to be approved!

```{r}

# Area saved / rejected proposals
n_reject <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>%
  filter(proposal_status=="Returned" | proposal_status=="Rejected" |
         proposal_status=="Closed" | proposal_status=="Rejected By REC" |
           proposal_status=="Withdrawn" | proposal_status=="CLOSED" |proposal_status=="REVOKED") %>% summarise() %>% tally(area_applied)%>%
  tally(n)

n_reject # 13,077.03 hectares of forests were not cleared post 2014

# Area cleared / approved proposals and/or pending
n_appro <- post_2014 %>% select(proposal_status,area_applied) %>%
  group_by(proposal_status, area_applied) %>% summarise() %>%   tally(area_applied) %>% tally(n) - n_reject

n_appro # 14,82,247 hectares of forests were cleared or approved to be cleared / pending

```

Let's calculate the area of forests approved or pending to be approved as function of the category they were binned under:

Mining alone contributed to maximum loss of forest area post 2014.

```{r}

# Here onwards we will use only the proposals that were approved or pending to be approved
appro_post_2014 <- post_2014 %>%
  filter(!(proposal_status=="Returned" | proposal_status=="Rejected" |
         proposal_status=="Closed" | proposal_status=="Rejected By REC" |
           proposal_status=="Withdrawn" | proposal_status=="CLOSED" |proposal_status=="REVOKED")) %>%
  filter(!(date=="NA" | date=="2008"))

# Data by category and area
cat_area <- appro_post_2014 %>% group_by(category) %>%
  summarise(Total_area = sum(area_applied))

# Create a plot
cat_plot <- ggplot(cat_area, aes(x = category , y = Total_area, fill= category)) +   
  geom_bar(stat="identity",color="black", size=0.08,position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+ theme_minimal() +
  labs(x="category", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none") +
  scico::scale_fill_scico_d(palette = "roma")

cat_plot

ggsave(cat_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\cat_area_post2014.png",
       width=12, height=7,dpi = 300)

```


Calculating area of forests approved to be cleared or pending to be cleared per year

See years 2016!

```{r}
# Data by year and area
appro_post_2014$date <- as.character(appro_post_2014$date)

area_year <- appro_post_2014 %>% group_by(date) %>%
  summarise(Total_area = sum(area_applied))

# Create a plot
area_year_plot <- ggplot(area_year, aes(x = date , y = Total_area)) +   
  geom_bar(stat="identity",color="black", size=0.08, position=position_dodge(), fill="blue",alpha=0.4) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+ theme_minimal() +
  labs(x="Year", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none")

area_year_plot

ggsave(area_year_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\year_area_post2014.png",
       width=12, height=7,dpi = 300)

```

Lets calculate the area of forests approved or pending to be approved as function of the state.  

```{r}
state_area <- appro_post_2014 %>% group_by(state_name) %>%
  summarise(Total_area = sum(area_applied))

# Create a plot
state_area_plot <- ggplot(state_area, aes(x = state_name , y = Total_area, fill=state_name)) +geom_bar(stat="identity",color="black", size=0.08,position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=3, vjust=-0.5)+
  theme_minimal() +
  labs(x="State", y="Total Area (in hectares)") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.ticks = element_blank(),
        legend.position = "none") +
  scico::scale_fill_scico_d(palette = "roma")

state_area_plot

ggsave(state_area_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\state_area_post2014.png",
       width=12, height=7,dpi = 300)

```


Lets calculate the area of forests approved or pending to be approved as function of the state and the category. 

Andhra Pradesh and Telangana are two states, that are seeing vast loss of forests since 2014

```{r}
# group by category and state
state_area <- appro_post_2014 %>% group_by(category, state_name) %>%
  summarise(Total_area=sum(area_applied))

nb.cols <- 24
mycolors <- colorRampPalette(brewer.pal(11, "Paired"))(nb.cols)

state_plot <- ggplot(state_area, aes(x = category , y = Total_area, fill= category)) +   
  geom_bar(stat="identity",color="black",size=0.08, position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=1.8, vjust=-0.5) + theme_minimal() +
  facet_wrap(~state_name) +
  labs(x="category", y="Total Area (in hectares)") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(size=8), 
        legend.text=element_text(size=8)) +
  scale_fill_manual(values = mycolors)
 
state_plot

ggsave(state_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\state_cat_area_post2014.png",
       width=20, height=10,dpi = 300)
```


Lets calculate the area of forests approved or pending to be approved as function of the category and year

Mining consistently increased until 2016

```{r}

cat_year <- appro_post_2014 %>% group_by(category, date) %>% 
  summarise(Total_area = sum(area_applied))

cat_year_plot <- ggplot(cat_year, aes(x = category , y = Total_area, fill= category)) + 
  geom_bar(stat="identity",color="black",size=0.08, position=position_dodge()) +
  geom_text(aes(label=round(Total_area)), position=position_dodge(width=1),size=1.8, vjust=-0.5) + theme_minimal() +
  facet_wrap(~date) +
  labs(x="Year", y="Total Area (in hectares)") + scale_fill_manual(values = mycolors) +
   theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(size=5), 
        legend.text=element_text(size=5))

cat_year_plot

ggsave(cat_year_plot, filename = "C:\\Users\\Vijay\\Desktop\\envt_clearance\\cat_year_post2014.png",
       width=15, height=10,dpi = 300)
  

```
